Adaptive Video Streaming with Optimized Bitstream Extraction and PID-based Quality Control

### 2015-05-19 tmm.tex
more problem analysis

In this section, we analyze the two most concerned problems in SVC adaptive video streaming, i.e., bitstream extraction and quality control, and introduce some existing related works on them.

@@

SVC bitstream extraction is substantially a combinatorial optimization problem. More specifically, it is in the form of ``0-1 Knapsack Problem" \cite{Knapsack}. In the ``0-1 Knapsack Problem", given a set of objects, each with a weight and a value, one decides which object to include in the knapsack so that the total weight is no more than a given limit and the total value is as large as possible. Similarly, in bitstream extraction problem, each NALU data packet can be treated as an item with certain weight (the packet's size) and value (its contribution to the video quality). The solving process is mainly to decide whether or not a packet should be included in the final extracted sub-stream, or in the term of ``0-1 Knapsack Problem", to choose which object to pack up.

Generally, the optimal solution of a typical ``0-1 Knapsack Problem" can be found using the dynamic programming algorithm. However, this is not operational for the bitstream extraction problem, mainly due to the dependence between packets. First, the value of a packet is not obvious and certain, since a packet's contribution to video quality may depend on the other packets. Second, the subset of packets to be extracted cannot be chosen arbitrarily, since if a packet is included, all the packets it depends on must also be included. In fact, to the best of our knowledge, no method presented in the literature claims to find the optimal solution of the SVC bitstream extraction problem. All the effort is made to obtain a solution that is as good as possible.

@@

For example, consider a video sequence decoded from a sub-stream which is extracted by discarding packet set $I_x$. The error between this sequence and the original sequence can be estimated as

	\begin{equation}
	\label{eq:subset_error}
e(I_x) = e_{full} + \sum_{i \in {I_x}} e_i \: .
	\end{equation}
Eq. (\ref{eq:subset_error}) has taken into consideration the error between the video sequence decoded from full enhancement packets and the original sequence, written as $e_{full}$.

@@

\begin{table*}[t]
\centering
\caption{Algorithm Performance: Luma PSNR Gain Compared with JSVM}
\label{tab:results}
\begin{minipage}{0.95\linewidth}
% \renewcommand{\thefootnote}{\thempfootnote}
\centering
\begin{tabular}{c|c|p{1.2cm}<{\centering}|p{1.2cm}<{\centering}|p{1.2cm}<{\centering}|p{1.2cm}<{\centering}| 
p{1.2cm}<{\centering}|p{1.2cm}<{\centering}|p{1.2cm}<{\centering}|p{1.2cm}<{\centering}}

%% |l|l| to left justify each column entry
%% |c|c| to center each column entry
%% use of \rule[]{}{} below opens up each row

\hline \hline
\multicolumn{2}{c|}{Sequence} &
{\em Bus} & {\em City} & {\em Crew} & {\em Football} & {\em Foreman} & {\em Harbour} & {\em Mobile} & {\em Soccer} \\ \hline

\multirow{2}{*}{JSVM QL\footnote{\label{footnote:JSVM_QL} Comparing the proposed algorithm with JSVM QL} (dB)}
& Max\footnote{\label{footnote:max} Maximum PSNR gain through all bitrate constraints} & 0.37 & 0.50 & 0.13 & 0.22 & 0.23 & 0.19 & 0.25 & 0.44 \\ \cline{2-10}
& Ave\footnote{\label{footnote:ave} Average PSNR gain through all bitrate constraints} & 0.11 & 0.11 & 0.05 & 0.12 & 0.05 & 0.05 & 0.06 & 0.13 \\ \hline

\multirow{2}{*}{JSVM no QL (dB)}
& Max & 0.50 & 0.83 & 0.32 & 0.29 & 0.50 & 0.34 & 0.61 & 0.53 \\ \cline{2-10}
& Ave & 0.18 & 0.37 & 0.20 & 0.15 & 0.22 & 0.15 & 0.25 & 0.29 \\ \hline

\end{tabular}
\end{minipage}
\end{table*}

@@

@@

\bibitem{Zhang12}
W.~Zhang, J.~Sun, J.~Liu and Z.~Guo, ``Optimized bit extraction of SVC exploiting linear error model,'' {\em 2012 IEEE International Symposium on Circuits and Systems}, pp.~1887--1890, Seoul, Korea, May 2012.

@@

\bibitem{Winken08}
M.~Winken, H.~Schwarz, T.~Wiegand, ``Joint Rate-distortion Optimization of Transform Coefficients for Spatial Scalable Video Coding Using SVC'' in {\em Proc. International Conference on Image Processing 2008}.

@@

\bibitem{Wong04}
C.W.Wong, O.C.Au and H.K.Lam, ``PID-based real-time rate control'', in {\em IEEE International Conference on Multimedia and Expo}, 1: 221-224, 2004.

>>>>

In this section, we analyze the two most concerned problems in SVC adaptive video streaming, i.e., bitstream extraction and quality control, and introduce some related works on them.

@@

Bitstream extraction is the operation of choosing a subset of data packets from a whole SVC bitstream to form a sub-stream. Fig. \ref{fig:Bitstream-Extraction} illustrates the structure of an SVC stream and a possible way to extract a sub-stream from it. The data packets in the stream are divided into different layers. The temporal layer (T0$\sim$T3) reflects the reference or dependency relationship between the frames, shown by the arrows on the top pointing from the frame which is referenced to the frame which is dependent on the reference. The quality layer (Q0$\sim$Q2) indicates scalability of video quality within each frame. The dash line demonstrates the sample extraction which will keep the data packets below the line and discard all the packets above it. This is only one option of the many possible ways, because there are many combinations of packets to be discarded. Our goal, however, is to choose a way that maximums the video quality of the extracted sub-stream, under the constrain of a give bitrate.

\begin{figure}[h]
\centering
\includegraphics[width = 0.9\linewidth]{Bitstream-Extraction.jpg}
\caption{Bitstream extraction of SVC \label{fig:Bitstream-Extraction}}
\end{figure}

SVC bitstream extraction, as we can see, is substantially a combinatorial optimization problem. More specifically, it is in the form of ``0-1 Knapsack Problem" \cite{Knapsack}. 
Generally, the optimal solution of a typical ``0-1 Knapsack Problem" can be found using the dynamic programming algorithm. However, this is not operational for the bitstream extraction problem, mainly due to the dependence between packets. First, the value of a packet is not obvious and certain, since a packet's contribution to video quality may depend on the other packets. Second, the subset of packets to be extracted cannot be chosen arbitrarily, since if a packet is included, all the packets it depends on must also be included. In fact, to the best of our knowledge, no method presented in the literature claims to find the optimal solution of the SVC bitstream extraction problem. All the effort is made to obtain a solution that is as good as possible.

@@

For example, consider a video sequence decoded from a sub-stream which is extracted by discarding packet set $I_x$. The error between this sequence and the original sequence can be estimated as
\begin{equation}
\label{eq:subset_error}
e(I_x) = e_{full} + \sum_{i \in {I_x}} e_i \: .
\end{equation}
Eq. (\ref{eq:subset_error}) has taken into consideration the error between the video sequence decoded from full enhancement packets and the original sequence, written as $e_{full}$.

@@

\begin{table*}[t]
\centering
\caption{Algorithm Performance: Luma PSNR Gain Compared with JSVM}
\label{tab:results}
\begin{minipage}{0.95\linewidth}
% \renewcommand{\thefootnote}{\thempfootnote}
\centering
\begin{tabular}{c|c|p{1.2cm}<{\centering}|p{1.2cm}<{\centering}|p{1.2cm}<{\centering}|p{1.2cm}<{\centering}| 
p{1.2cm}<{\centering}|p{1.2cm}<{\centering}|p{1.2cm}<{\centering}|p{1.2cm}<{\centering}}
%% |l|l| to left justify each column entry
%% |c|c| to center each column entry
%% use of \rule[]{}{} below opens up each row 
\hline \hline
\multicolumn{2}{c|}{Sequence} &
{\em Bus} & {\em City} & {\em Crew} & {\em Football} & {\em Foreman} & {\em Harbour} & {\em Mobile} & {\em Soccer} \\ \hline 
\multirow{2}{*}{JSVM QL\footnote{\label{footnote:JSVM_QL} Comparing the proposed algorithm with JSVM QL} (dB)}
& Max\footnote{\label{footnote:max} Maximum PSNR gain through all bitrate constraints} & 0.37 & 0.50 & 0.13 & 0.22 & 0.23 & 0.19 & 0.25 & 0.44 \\ \cline{2-10}
& Ave\footnote{\label{footnote:ave} Average PSNR gain through all bitrate constraints} & 0.11 & 0.11 & 0.05 & 0.12 & 0.05 & 0.05 & 0.06 & 0.13 \\ \hline
\multirow{2}{*}{JSVM no QL (dB)}
& Max & 0.50 & 0.83 & 0.32 & 0.29 & 0.50 & 0.34 & 0.61 & 0.53 \\ \cline{2-10}
& Ave & 0.18 & 0.37 & 0.20 & 0.15 & 0.22 & 0.15 & 0.25 & 0.29 \\ \hline
\end{tabular}
\end{minipage}
\end{table*}

@@

\bibitem{Ramanathan12}
P.~Ramanathan and N.~Jayant. ``RD optimized bitstream extraction for H. 264/SVC based video streaming'' In {\em Proceedings of IEEE International Conference on Image Processing}, pp. 2241-2244, 2012.

\bibitem{Yang13}
K.~Yang, S.~Wan, Y.~Gong, and Y.~Feng. ``A fast algorithm of bitstream extraction using distortion prediction based on simulated annealing'' In {\em Journal of Visual Communication and Image Representation}, no.7, 752-759, 2013.

@@

\bibitem{Zhang12}
W.~Zhang, J.~Sun, J.~Liu and Z.~Guo, ``Optimized bit extraction of SVC exploiting linear error model,'' {\em Proceedings of IEEE International Symposium on Circuits and Systems}, pp.~1887--1890, Seoul, Korea, May 2012.

@@

\bibitem{Winken08}
M.~Winken, H.~Schwarz, T.~Wiegand, ``Joint Rate-distortion Optimization of Transform Coefficients for Spatial Scalable Video Coding Using SVC'' in {\em Proceedings of IEEE International Conference on Image Processing}, 2008.

@@

\bibitem{Wong04}
C.W.Wong, O.C.Au and H.K.Lam, ``PID-based real-time rate control'', in {\em Proceedings of IEEE International Conference on Multimedia and Expo}, 1: 221-224, 2004.
